---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
autor: Micha³ Bukowski
---

## Loading and preprocessing the data

```{r}
#loading data.table package
library(data.table)

#loading the data
activity <- data.table(
                read.csv(
                        file="activity/activity.csv"
                        , colClasses=c("numeric", "character", "numeric")
                        , na.strings="NA"                        
                )
        )
#converting date column to Date class
activity[, date := as.Date(date)]

summary(activity)
```

## What is mean total number of steps taken per day?
```{r, fig.width=12}
library(ggplot2)

#generating a histogram of total number of steps taken each day
g <- ggplot(activity[!is.na(steps)], aes(x=factor(date), y=steps)) +
        geom_bar(stat="identity") +
        ggtitle("Total number of steps taken each day") +
        labs(x="Date", y = "Number of steps") +
        theme(axis.text.x = element_text(angle=90,vjust=0.5,hjust=1)) 
print(g)  
```
```{r computetime}
#creating table containing sum of steps per day
sum.of.steps.by.date <- activity[
                ,list(sum.of.steps = sum(steps))
                ,by=date
        ]

#making variables to use in the text below
steps.mean <- mean(sum.of.steps.by.date$sum.of.steps, na.rm=TRUE)
steps.median <- median(sum.of.steps.by.date$sum.of.steps, na.rm=TRUE)
```
`r steps.mean` is the mean total number of steps taken per day.

`r steps.median` is the median total number of step taken per day.


## What is the average daily activity pattern?
```{r, fig.width=12}
#creating table containing average number of steps per interval
avg.steps.by.interval <- activity[
                ,list(avg.steps = mean(steps, na.rm=TRUE))
                ,by=interval
        ]

#generating a time series plot of average number of steps taken per interval
g <- ggplot(avg.steps.by.interval, aes(interval, avg.steps)) +
        geom_line() +
        ggtitle("Average number of steps taken per interval") +
        labs(x="Interval", y = "Number of steps") +
        theme_bw() 
print(g)
```
```{r}
#calculating the interval with the maximum average number of steps
max.avg.steps <- avg.steps.by.interval[,max:=max(avg.steps)][avg.steps==max]

```
The average number of steps across all days occures in the `r max.avg.steps$interval` interval and amounts to `r max.avg.steps$max` average number of steps.

## Imputing missing values
```{r}
#calculating number of missing values
sum.of.na.rows <- sum(activity[,is.na(steps)])
```
Number of rows with missing value of steps amounts to `r sum.of.na.rows`.

```{r}
#
#In this chunk I would prefer to write echo=FALSE but it was forbiden to do this :)
#

#setting english setlocale
local.time <- Sys.getlocale("LC_TIME")
Sys.setlocale("LC_TIME", "C")

#filling missing values with the average number of steps per interval per weekday
filled.activity <- activity
filled.activity[
                ,week.day := weekdays(date)
        ][
                ,`:=` (avg.steps = mean(steps, na.rm=TRUE))
                ,by=c("interval", "week.day")
        ]
filled.activity[
        !is.na(steps)
        ,`:=`(filled.steps=steps)
]
filled.activity[
        is.na(steps)
        ,`:=`(filled.steps=avg.steps)
]

#setting previous datime settings
Sys.setlocale("LC_TIME", local.time)

```
```{r, fig.width=12}
#generating a histogram of total number of steps taken each day for the filled data
g <- ggplot(filled.activity, aes(x=factor(date), y=filled.steps)) +
        geom_bar(stat="identity") +
        ggtitle("Total number of steps taken each day") +
        labs(x="Date", y = "Number of steps") +
        theme(axis.text.x = element_text(angle=90,vjust=0.5,hjust=1)) 
print(g)  
```
```{r}
#creating table containing sum of steps per day with the filled data
sum.of.steps.by.date <- filled.activity[
                ,list(sum.of.steps = sum(filled.steps))
                ,by=date
        ]

#making variables to use in the text below
steps.mean <- mean(sum.of.steps.by.date$sum.of.steps)
steps.median <- median(sum.of.steps.by.date$sum.of.steps)
```
`r steps.mean` is the mean total number of steps taken per day considering filled data.  
`r steps.median` is the median total number of step taken per day considering filled data.  
As you can observe, both mean and median number of steps in the filled data has growth a little.
## Are there differences in activity patterns between weekdays and weekends?
```{r}
#creating a variable indicating if it's a weekday or weekend
filled.activity[
        week.day == c("Sunday","Saterday")
        ,`:=`(day.type="weekend")
]
filled.activity[
        week.day != c("Sunday","Saterday")
        ,`:=`(day.type="weekday")
]
filled.activity[,day.type:=factor(day.type)]
```
```{r, fig.width=12}
#creating table containing average number of steps per interval and type of day
avg.steps.by.interval <- filled.activity[
                ,list(avg.steps = mean(filled.steps))
                ,by=c("interval", "day.type")
        ]

#generating a time series plot of average number of steps taken per interval and type of day
g <- ggplot(avg.steps.by.interval, aes(interval, avg.steps)) +
        facet_grid(day.type ~ .) +
        geom_line() +
        ggtitle("Average number of steps taken per interval and type of day") +
        labs(x="Interval", y = "Number of steps") +
        theme_bw() 
print(g)
```
